On the uniqueness and stability of dictionaries for sparse representation of noisy signals

### 2018-12-24 manuscripts/IEEE/chazthm_ieee_trans_sig_CHAZ.tex
new intro

\begin{abstract}
Dictionary learning for sparse linear coding has exposed characteristic properties of natural signals.
However, a universal theorem guaranteeing the consistency of estimation in this model is lacking.
Here, we prove that for all diverse enough datasets generated from the sparse coding model, latent dictionaries and codes are uniquely and stably determined up to measurement error.  Applications are given to data analysis, engineering, and neuroscience. 
\end{abstract}

@@

\IEEEPARstart{B}{lind} source separation is a classical problem in signal processing \cite{sato1975method}.
In one common modern formulation, each of $N$ observed $n$-dimensional signals is a (noisy) linear combination of at most $k$ elementary waveforms drawn from some unknown dictionary of size $m$, typically with $m \ll N$ (see \cite{Zhang15} for a comprehensive review of this and related models).
Approximating solutions to this sparsity-constrained inverse problem have provided insight into the structure of many signal classes lacking domain-specific formal models (e.g., in vision \cite{wang2015sparse}).  In particular, it has been shown that response properties of simple-cell neurons in mammalian visual cortex emerge from optimizing a dictionary to represent small patches of natural images \cite{Olshausen96, hurri1996image, bell1997independent, van1998independent}, a major advance in computational neuroscience. A curious aspect of this finding is that the latent waveforms (e.g., ``Gabor'' wavelets) estimated from data appear to be canonical \cite{donoho2001can};
i.e., they are found in learned dictionaries independent of algorithm or image training set.

Motivated by these discoveries and earlier work in the theory of neural communication \cite{Isely10}, we address when dictionaries and the sparse representations they induce are uniquely determined by data.  Answers to this question also have other real-world implications.  For example, a sparse coding analysis of local painting style can be used for forgery detection \cite{hughes2010}, but only if all dictionaries consistent with training data do not differ appreciably in their ability to sparsely encode new samples. 
Fortunately, algorithms with proven recovery of generating dictionaries under certain 
conditions have recently been proposed (see \cite[Sec.~I-E]{Sun16} for a summary of the state-of-the-art). Few theorems, however, can be cited to explain this uniqueness more broadly; in particular, a universal guarantee in the context of noise has yet to emerge.

Here, we prove very generally that uniqueness and stability is an expected property of the sparse linear coding model. 
More specifically, dictionaries that preserve sparse codes (e.g., satisfy a ``spark condition'') are identifiable from as few as \mbox{$N = m(k-1){m \choose k} + m$} noisy sparse linear combinations of their columns up to an error that is linear in the noise (Thm.~\ref{DeterministicUniquenessTheorem}). In fact, provided $n \geq \min(2k,m)$, in almost all cases the dictionary learning problem is well-posed (as per Hadamard \cite{Hadamard1902}) given enough data (Cor.~\ref{ProbabilisticCor}). Moreover, these explicit, algorithm-independent guarantees hold without assuming the recovered matrix satisfies a spark condition, and even when the number $m$ of dictionary elements is unknown.

More formally, let $\mathbf{A} \in \mathbb R^{n \times m}$ be a matrix with columns $\mathbf{A}_j$ ($j = 1,\ldots,m$) and let dataset $Z$ consist of measurements:
\begin{align}\label{LinearModel}
\mathbf{z}_i = \mathbf{A}\mathbf{x}_i + \mathbf{n}_i,\ \ \  \text{$i=1,\ldots,N$},
\end{align}
for $k$-\emph{sparse} $\mathbf{x}_i \in \mathbb{R}^m$ having at most $k<m$ nonzero entries and \emph{noise} $\mathbf{n}_i \in \mathbb{R}^n$, with bounded norm $\| \mathbf{n}_i \|_2 \leq  \eta$ representing our combined worst-case uncertainty in  measuring $\mathbf{A}\mathbf{x}_i$.
The first mathematical problem we address is the following.

\begin{problem}[Sparse linear coding]\label{InverseProblem}
Find a matrix $\mathbf{B}$ and $k$-sparse $\mathbf{\overline x}_1, \ldots, \mathbf{\overline x}_N \in \mathbb{R}^{\overline m}$ with $\|\mathbf{z}_i - \mathbf{B}\mathbf{\overline x}_i\|_2 \leq \eta$ for all $i$.
\end{problem}

Note that any particular solution $(\mathbf{B}; \mathbf{\overline x}_1 \ldots, \mathbf{\overline x}_N)$ to this problem gives rise to an orbit of equivalent solutions $(\mathbf{BPD}; \mathbf{D}^{-1}\mathbf{P}^{\top}\mathbf{\overline x}_1, \ldots, \mathbf{D}^{-1}\mathbf{P}^{\top}\mathbf{\overline x}_N)$, where $\mathbf{P}$ is any permutation matrix and $\mathbf{D}$ any invertible diagonal. Previous theoretical work addressing the noiseless case $\eta =0$ (e.g., \cite{li2004analysis, Georgiev05, Aharon06, Hillar15}) with $\overline m = m$ has shown that a solution to Prob.~\ref{InverseProblem} (when it exists) is unique up to this  ambiguity provided the $\mathbf{x}_i$ are sufficiently diverse and the matrix $\mathbf{A}$ satisfies the \textit{spark condition}:
\begin{align}\label{SparkCondition}
\mathbf{A}\mathbf{x}_1 = \mathbf{A}\mathbf{x}_2 \implies \mathbf{x}_1 = \mathbf{x}_2, \ \ \ \text{for all $k$-sparse } \mathbf{x}_1, \mathbf{x}_2,
\end{align}
%
which would be, in any case, necessary for uniqueness of the $\mathbf{x}_i$. Here, we study stability in the practical setting of noise.

@@

To see how Def. \ref{maindef} directly relates to Prob. \ref{InverseProblem}, suppose that $Y$ has a stable $k$-sparse representation in $\mathbb{R}^m$ and fix $\delta_1, \delta_2$ to be the desired recovery accuracy in \eqref{def1}. Consider any dataset $Z$ generated as in \eqref{LinearModel} with $\eta \leq \frac{1}{2} \varepsilon(\delta_1, \delta_2)$. Then from the triangle inequality, it follows that any matrix $\mathbf{B}$ and $k$-sparse $\mathbf{\overline x}_1, \ldots, \mathbf{\overline x}_N \in \mathbb{R}^m$ solving Prob.~\ref{InverseProblem} are necessarily within $\delta_1, \delta_2$ of the original dictionary $\mathbf{A}$ and codes $\mathbf{x}_i$, respectively.

In the next section, we give precise statements of our main results, which include an explicit form for $\varepsilon(\delta_1, \delta_2)$. We then prove our main theorem (Thm.~\ref{DeterministicUniquenessTheorem}) in Sec.~\ref{DUT} after stating some additional definitions and lemmas required for the proof, including a useful result in combinatorial matrix analysis (Lem.~\ref{MainLemma}, proven in Appendix). We also provide an argument that extends our guarantees to the following more common optimization formulation of the dictionary learning problem (Thm.~\ref{SLCopt}).

@@

%We delay defining the explicit constant $C_1$ until Section \ref{DUT} (\eqref{Cdef1}).
%To be clear, the implication of Thm.~\ref{DeterministicUniquenessTheorem} is that $Y = \{\mathbf{Ax}_1, \ldots, \mathbf{Ax}_N\}$ has a stable $k$-sparse representation in $\mathbb{R}^m$, with \eqref{def1} guaranteed provided $\varepsilon$ in Def.~\ref{maindef} does not exceed: 
In words, Thm.~\ref{DeterministicUniquenessTheorem} says that the smaller the difference $\overline m - m$ between recovery and original latent dictionary sizes, the more columns and coefficients of the original dictionary $\mathbf{A}$ and codes $\mathbf{x}_i$ are contained (up to noise) in the appropriately scaled dictionary $\mathbf{B}$ and codes $\mathbf{\overline x}_i$.  In the particular case when $\overline m = m$, the theorem directly implies that  $Y = \{\mathbf{Ax}_1, \ldots, \mathbf{Ax}_N\}$ has a stable $k$-sparse representation in $\mathbb{R}^m$, with inequalities \eqref{def1} guaranteed for $\varepsilon$ in Def.~\ref{maindef} given by: 
\begin{align}\label{epsdel}
\varepsilon(\delta_1, \delta_2) := \min \left\{ \frac{\delta_1}{ C_1 }, \frac{ \delta_2 L_{2k}(\mathbf{A})}{ 1 + C_1 \left( \delta_2 + \max_{i \in [N]} \|\mathbf{x}_i\|_1  \right) } \right\}.
\end{align}

Note that sparse codes $\mathbf{x}_i$ with a shared support that are in general linear position are straightforward to produce using a ``Vandermonde'' matrix construction (i.e., use the columns of the matrix $[\gamma_{i}^j]_{i,j=1}^{k,N}$, for distinct nonzero $\gamma_i$).  Thus, the assumptions of Thm.~\ref{DeterministicUniquenessTheorem} are easily met, leading to the following direct application to uniqueness in sparse linear coding. % (Prob.~\ref{InverseProblem}).
%=======
%An immediate practical implication of this result is that there exists a practical procedure to affirm if one's proposed solution $(\mathbf{B}, \mathbf{\overline x}_1, \ldots, \mathbf{x}_N)$ to Prob.~\ref{InverseProblem} is indeed unique (up to noise and inherent ambiguities): simply check that $\mathbf{B}$ and the $\mathbf{\overline x}_i$ satisfy the assumptions on $\mathbf{A}$ and the $\mathbf{x}_i$ in Thm.~\ref{DeterministicUniquenessTheorem}.
%
%%In fact, a more general result (stated clearly in the next section) can be gleaned from our method of proving Thm.~\ref{DeterministicUniquenessTheorem}. Briefly, in cases where $\mathbf{B}$ has $\overline m \neq m$ columns, or $\mathcal{H}$ is not regular or only partially satisfying the SIP, a relation between $\overline m$ and the degree sequence of nodes in $\mathcal{H}$ gives indices $J \subseteq [m]$ defining a submatrix $\mathbf{A}_J$ and subvectors $\mathbf{x}_i^J$ that are recoverable in the sense of \eqref{Cstable} and \eqref{b-PDa}. For example, if $\mathcal{H}$ is $\ell$-regular with the SIP but $m \leq \overline m < m\ell/(\ell - 1)$ then we have nonzero $|J| = \overline m - \ell(\overline m - m)$. The implication here is that the smaller the difference $\overline m - m$, the more columns and code entries of the original $n \times m$ dictionary $\mathbf{A}$ and codes $\mathbf{x}_i$ contained (up to noise) in the appropriately scaled $n \times \overline m$ dictionary $\mathbf{B}$ and codes $\mathbf{\overline x}_i$. When $\overline m = m$, we recover Thm.~\ref{DeterministicUniquenessTheorem}.
%
%%In fact, even if $\mathcal{H}$ is not regular or only partially satisfies the SIP, a relation between $\overline m$ and the degree sequence of nodes in $\mathcal{H}$ may give the indices $J \subseteq [m]$. For sake of brevity, we delay to the next section a clear statement of this more general result.
%
%Regarding the assumptions of Thm.~\ref{DeterministicUniquenessTheorem}, it so happens that sparse codes $\mathbf{x}_i$ in general linear position are straightforward to produce with a ``Vandermonde'' matrix construction \cite{Hillar15}, leading to the following.
%>>>>>>> d1407a275dc391dabad7bb62d3e659f9a4ac4624

@@

There are other less direct consequences of Thm.~\ref{DeterministicUniquenessTheorem}.  For instance, we use it to prove uniqueness for the optimization formulation of sparse linear coding, Prob.~\ref{SLCopt}, the main object of interest for those applying dictionary learning to their data.

@@

\begin{proposition}\label{neuroprop}
Sparse neural population activity is recoverable from noisy random linear compression by any method that solves Prob.~\ref{DeterministicUniquenessTheorem} or \ref{SLCopt}; in particular, via biophysically plausible unsupervised sparse linear coding (e.g., \cite{rehnsommer2007, rozell2007neurally, pehlevan2015normative}).
\end{proposition}
%\cite{rehnsommer2007, rozell2007neurally, ganguli2012compressed, hu2014hebbian}

We close this section with comments on optimality.  Our linear scaling for $\varepsilon$ in \eqref{epsdel} is essentially optimal (e.g., see \cite{arias2013fundamental}), but a basic open problem remains: how many samples are necessary to determine the sparse linear coding model? If $k$ is held fixed or if the size of the support set of reconstructing codes is known to be polynomial in $\overline m$ and $k$, then a practical (polynomial) amount of data suffices.\footnote{In the latter case, a reexamination of the pigeonholing argument in the proof of Thm.~\ref{DeterministicUniquenessTheorem} requires a polynomial number of samples distributed over a polynomial number of supports.} Reasons to be skeptical that this holds in general, however, can be found in \cite{tillmann2014computational, Tillmann15}.

@@

The goal of this work was to explain the emergence of characteristic representations from sparse linear coding models fit to natural data, despite the varied assumptions underlying the many algorithms in current use. To this end, we have taken an important step toward unifying the great number of publications on the topic by demonstrating very general, deterministic conditions under which identification of parameters in this parsimonious model is not only possible but also robust to the inevitable uncertainty permeating measurement and model choice.

@@

An immediate application of our theoretical work is Prop.~\ref{neuroprop}, which certifies the validity of a theory of neurally-plausible bottleneck communication in the brain: that sparse linear coding in a compressed space of neural activity can recover sparse codes sent through a randomly-constructed (but unknown) noisy wiring bottleneck \cite{Isely10}.\footnote{We refer the reader to \cite{ganguli2012compressed} for more on interactions between dictionary learning and neuroscience.}

@@

In fact, uniqueness guarantees with minimal assumptions apply to all areas of data science and engineering that utilize learned sparse structure. For example, several groups have applied compressed sensing to signal processing tasks: MRI analysis \cite{lustig2008compressed}, image compression \cite{Duarte08}, and, more recently, the design of an ultrafast camera \cite{Gao14}. Given such effective uses of compressed sensing, it is only a matter of time before these systems incorporate dictionary learning to encode and decode signals (e.g., in a device that learns structure from motion \cite{kong2016prior}), just as scientists have used it to make sense of their data \cite{jung2001imaging, Agarwal14, lee2016sparse, wu2016stability}. Assurances such as those offered by our theorems certify that different devices (with different initialization, etc.) will learn equivalent representations given enough data from statistically identical systems.\footnote{To contrast with the current hot topic of ``Deep Learning'', there are few such uniqueness guarantees for these models of data; moreover, even small noise can dramatically alter their output \cite{goodfellow2014explaining}.} Indeed, it seems a main reason for the sustained interest in dictionary learning as an unsupervised method for data analysis is the assumed well-posedness of parameter identification in the sparse linear coding model, confirmation of which forms the core of our theoretical findings.

@@

%Since its inception some twenty years ago, sparse linear coding has become a standard tool in signal analysis, yielding myriad insights into the structure of natural signals across a variety of scientific domains. In this work,

>>>>

\begin{abstract}
Learning optimal dictionaries for sparse coding has exposed characteristic sparse features of many natural signals. However, universal conditions guaranteeing the consistency of these learned features in the presence of measurement noise have yet to be stated. Here, we prove very generally that optimal dictionaries and sparse codes are uniquely and stably determined up to measurement error for all diverse enough datasets generated in accordance with the constraints assumed by sparse coding. Applications are given to data analysis, engineering, and neuroscience. 
\end{abstract}

@@

% \cite{sato1975method}
\IEEEPARstart{A}{}common assumption underlying contemporary applications of signal processing and pattern analysis is that each of $N$ observed $n$-dimensional signal samples is a (noisy) linear combination of at most $k$ elementary waveforms drawn from some unknown \emph{dictionary} of size $m$, typically with $m \ll N$ (see \cite{Zhang15} for a comprehensive review of this and related models). Attempts to approximate solutions to this sparsity-constrained inverse problem have provided insight into the structure of many signal classes lacking idiosyncratic formal models (e.g., in vision \cite{wang2015sparse}). In fact, a seminal work in the field demonstrated that dictionaries optimized to sparsely represent small patches of natural images share many qualitative similarities to filters fit to the response properties of simple-cell neurons in mammalian visual cortex \cite{Olshausen96, hurri1996image, bell1997independent, van1998independent}, a major result in theoretical neuroscience. Moreover, these latent waveforms (e.g., ``Gabor'' wavelets) estimated from data appear to be canonical \cite{donoho2001can}; i.e. they emerge in dictionaries learned by a variety of algorithms trained with respect to different natural image datasets.

Motivated by these discoveries and earlier work proposing a related theory of communication between sparsely active neural populations \cite{Isely10}, we address when dictionaries and the sparse representations they induce are uniquely determined by data. Answers to this question have real-world implications. For example, it has been demonstrated that a sparse coding analysis of painting style can detect forgeries \cite{hughes2010}; but this implies all dictionaries consistent with training data generalize in their ability to sparsely encode new samples. Fortunately, several algorithms have recently been proposed which provably recover latent dictionaries under idiosyncratic conditions (see \cite[Sec.~I-E]{Sun16} for a summary of the state-of-the-art). Few theorems, however, can be cited to explain the consistency of inference with respect to this model more broadly; in particular, a universal guarantee of uniqueness and stability in the context of noise has yet to emerge.

Here, we prove very generally that uniqueness and stability of solutions is an expected property of both the sparse coding problem described above as well as a related (and perhaps more well-known) optimization formulation, which instead asks for a dictionary and sparse codes that minimize the mean sparsity over samples.

Formally, let $\mathbf{A} \in \mathbb R^{n \times m}$ be a matrix with columns $\mathbf{A}_j$ ($j = 1,\ldots,m$) and let dataset $Z$ consist of measurements:
\begin{align}\label{LinearModel}
\mathbf{z}_i = \mathbf{A}\mathbf{x}_i + \mathbf{n}_i,\ \ \  \text{$i=1,\ldots,N$},
\end{align}
for $k$-\emph{sparse} $\mathbf{x}_i \in \mathbb{R}^m$ having at most $k<m$ nonzero entries and \emph{noise} $\mathbf{n}_i \in \mathbb{R}^n$, with bounded norm $\| \mathbf{n}_i \|_2 \leq  \eta$ representing our combined worst-case uncertainty in  measuring $\mathbf{A}\mathbf{x}_i$. We show that dictionaries $\mathbf{A}$ injective on certain subsets of $k$-sparse codes are identifiable from as few as \mbox{$N = m(k-1){m \choose k} + m$} noisy sparse linear combinations of their columns up to an error that is linear in the noise (Thm.~\ref{DeterministicUniquenessTheorem}). Specifically, the first mathematical problem we consider is:

\begin{problem}\label{InverseProblem}
Find a matrix $\mathbf{B}$ and $k$-sparse $\mathbf{\overline x}_1, \ldots, \mathbf{\overline x}_N \in \mathbb{R}^{\overline m}$ satisfying $\|\mathbf{z}_i - \mathbf{B}\mathbf{\overline x}_i\|_2 \leq \eta$ for all $i$.
\end{problem}

Our claim is that there exist very general conditions under which any matrix $\mathbf{B}$ that solves this problem is always ``close" to $\mathbf{A}$, and likewise for the sparse codes $\mathbf{\overline x}_i$ and $\mathbf{x}_i$. In fact, provided $n \geq \min(2k,m)$, in almost all cases \ref{InverseProblem} is well-posed (as per Hadamard \cite{Hadamard1902}) given enough data (Cor.~\ref{ProbabilisticCor}). In particular, these explicit, universal guarantees hold without imposing \emph{any} constraints on the recovered matrix $\mathbf{B}$ (e.g. that it, too, satisfy any injectivity criteria), even the number $m$ of dictionary elements it should contain.

Note that any particular solution to Problem \ref{InverseProblem} represents an infinity of equivalent solutions $\mathbf{BPD}$ and $\mathbf{D}^{-1}\mathbf{P}^{\top}\mathbf{\overline x}_1, \ldots, \mathbf{D}^{-1}\mathbf{P}^{\top}\mathbf{\overline x}_N$ parametrized by a choice of permutation matrix $\mathbf{P}$ and invertible diagonal matrix $\mathbf{D}$ capturing the sparsity-preserving labeling and scaling ambiguities inherent to the problem statement. Previous theoretical work addressing the noiseless case $\eta =0$ (e.g., \cite{li2004analysis, Georgiev05, Aharon06, Hillar15}) with fixed $\overline m = m$ has shown that the solution to Prob.~\ref{InverseProblem}, when it exists, is unique up to these  ambiguities provided the $\mathbf{x}_i$ are sufficiently diverse and the matrix $\mathbf{A}$ satisfies:
\begin{align}\label{SparkCondition}
\mathbf{A}\mathbf{x}_1 = \mathbf{A}\mathbf{x}_2 \implies \mathbf{x}_1 = \mathbf{x}_2, \ \ \ \text{for all $k$-sparse } \mathbf{x}_1, \mathbf{x}_2,
\end{align}
%
also called the \textit{spark condition}, which in any case is necessary to guarantee the uniqueness of arbitrary sequences of $k$-sparse $\mathbf{x}_i$. We generalize these results to the practical setting of additive noise by incorporating the notion of stability with respect to noise.

@@

To see how Def. \ref{maindef} is motivated by Prob. \ref{InverseProblem}, suppose that $Y$ has a stable $k$-sparse representation in $\mathbb{R}^m$ and fix $\delta_1, \delta_2$ to be the desired recovery accuracy in \eqref{def1}. Now, consider any dataset $Z$ generated as in \eqref{LinearModel} with $\eta \leq \frac{1}{2} \varepsilon(\delta_1, \delta_2)$. From the triangle inequality, it follows that any matrix $\mathbf{B}$ and $k$-sparse $\mathbf{\overline x}_1, \ldots, \mathbf{\overline x}_N \in \mathbb{R}^m$ solving Prob.~\ref{InverseProblem} are necessarily within $\delta_1, \delta_2$ of the original dictionary $\mathbf{A}$ and codes $\mathbf{x}_i$, respectively.

In the next section, we give precise statements of our main results, which include an explicit form for $\varepsilon(\delta_1, \delta_2)$. We then prove our main theorem (Thm.~\ref{DeterministicUniquenessTheorem}) in Sec.~\ref{DUT} after stating some additional definitions and lemmas required for the proof, including a useful result in combinatorial matrix analysis (Lem.~\ref{MainLemma}, proven in Appendix). We also provide a short argument extending our guarantees to the following more common formulation of the dictionary learning problem as an optimization over model parameters minimizing the total (or equivalently, the mean) number of non-zero entries in the sparse codes (Thm.~\ref{SLCopt}):

@@

%We delay defining the explicit constant $C_1$ until Section \ref{DUT} (\eqref{Cdef1}).
%To be clear, the implication of Thm.~\ref{DeterministicUniquenessTheorem} is that $Y = \{\mathbf{Ax}_1, \ldots, \mathbf{Ax}_N\}$ has a stable $k$-sparse representation in $\mathbb{R}^m$, with \eqref{def1} guaranteed provided $\varepsilon$ in Def.~\ref{maindef} does not exceed: 
In words, Thm.~\ref{DeterministicUniquenessTheorem} says that the smaller the difference $\overline m - m$ between assumed and actual latent dictionary sizes, the more columns and coefficients of the original dictionary $\mathbf{A}$ and codes $\mathbf{x}_i$ are contained (up to noise) in the appropriately scaled dictionary $\mathbf{B}$ and codes $\mathbf{\overline x}_i$.  In the particular case when $\overline m = m$, the theorem directly implies that  $Y = \{\mathbf{Ax}_1, \ldots, \mathbf{Ax}_N\}$ has a stable $k$-sparse representation in $\mathbb{R}^m$, with inequalities \eqref{def1} guaranteed for $\varepsilon$ in Def.~\ref{maindef} given by: 
\begin{align}\label{epsdel}
\varepsilon(\delta_1, \delta_2) := \min \left\{ \frac{\delta_1}{ C_1 }, \frac{ \delta_2 L_{2k}(\mathbf{A})}{ 1 + C_1 \left( \delta_2 + \max_{i \in [N]} \|\mathbf{x}_i\|_1  \right) } \right\}.
\end{align}

Note that sparse codes $\mathbf{x}_i$ with a shared support that are in general linear position are straightforward to produce using a ``Vandermonde'' matrix construction (i.e., use the columns of the matrix $[\gamma_{i}^j]_{i,j=1}^{k,N}$, for distinct nonzero $\gamma_i$).  Thus, the assumptions of Thm.~\ref{DeterministicUniquenessTheorem} are easily met, leading to the following direct application to uniqueness in sparse coding. % (Prob.~\ref{InverseProblem}).
%=======
%An immediate practical implication of this result is that there exists a practical procedure to affirm if one's proposed solution $(\mathbf{B}, \mathbf{\overline x}_1, \ldots, \mathbf{x}_N)$ to Prob.~\ref{InverseProblem} is indeed unique (up to noise and inherent ambiguities): simply check that $\mathbf{B}$ and the $\mathbf{\overline x}_i$ satisfy the assumptions on $\mathbf{A}$ and the $\mathbf{x}_i$ in Thm.~\ref{DeterministicUniquenessTheorem}.
%
%%In fact, a more general result (stated clearly in the next section) can be gleaned from our method of proving Thm.~\ref{DeterministicUniquenessTheorem}. Briefly, in cases where $\mathbf{B}$ has $\overline m \neq m$ columns, or $\mathcal{H}$ is not regular or only partially satisfying the SIP, a relation between $\overline m$ and the degree sequence of nodes in $\mathcal{H}$ gives indices $J \subseteq [m]$ defining a submatrix $\mathbf{A}_J$ and subvectors $\mathbf{x}_i^J$ that are recoverable in the sense of \eqref{Cstable} and \eqref{b-PDa}. For example, if $\mathcal{H}$ is $\ell$-regular with the SIP but $m \leq \overline m < m\ell/(\ell - 1)$ then we have nonzero $|J| = \overline m - \ell(\overline m - m)$. The implication here is that the smaller the difference $\overline m - m$, the more columns and code entries of the original $n \times m$ dictionary $\mathbf{A}$ and codes $\mathbf{x}_i$ contained (up to noise) in the appropriately scaled $n \times \overline m$ dictionary $\mathbf{B}$ and codes $\mathbf{\overline x}_i$. When $\overline m = m$, we recover Thm.~\ref{DeterministicUniquenessTheorem}.
%
%%In fact, even if $\mathcal{H}$ is not regular or only partially satisfies the SIP, a relation between $\overline m$ and the degree sequence of nodes in $\mathcal{H}$ may give the indices $J \subseteq [m]$. For sake of brevity, we delay to the next section a clear statement of this more general result.
%
%Regarding the assumptions of Thm.~\ref{DeterministicUniquenessTheorem}, it so happens that sparse codes $\mathbf{x}_i$ in general linear position are straightforward to produce with a ``Vandermonde'' matrix construction \cite{Hillar15}, leading to the following.
%>>>>>>> d1407a275dc391dabad7bb62d3e659f9a4ac4624

@@

There are other less direct consequences of Thm.~\ref{DeterministicUniquenessTheorem}.  For instance, we apply it to prove uniqueness and stability of solutions to Prob.~\ref{SLCopt}, the usual optimization problem of interest for those applying dictionary learning to their data.

@@

\begin{proposition}\label{neuroprop}
Sparse neural population activity is recoverable from noisy random linear compression by any method that solves Prob.~\ref{DeterministicUniquenessTheorem} or \ref{SLCopt}; in particular, via biophysically plausible unsupervised sparse coding (e.g., \cite{rehnsommer2007, rozell2007neurally, pehlevan2015normative}).
\end{proposition}
%\cite{rehnsommer2007, rozell2007neurally, ganguli2012compressed, hu2014hebbian}

We close this section with comments on optimality.  Our linear scaling for $\varepsilon$ in \eqref{epsdel} is essentially optimal (e.g., see \cite{arias2013fundamental}), but a basic open problem remains: how many samples are necessary to determine the sparse coding model? If $k$ is held fixed or if the size of the support set of reconstructing codes is known to be polynomial in $\overline m$ and $k$, then a practical (polynomial) amount of data suffices.\footnote{In the latter case, a reexamination of the pigeonholing argument in the proof of Thm.~\ref{DeterministicUniquenessTheorem} requires a polynomial number of samples distributed over a polynomial number of supports.} Reasons to be skeptical that this holds in general, however, can be found in \cite{tillmann2014computational, Tillmann15}.

@@

The goal of this work was to explain the emergence of characteristic representations from sparse coding models fit to natural data, despite the varied assumptions underlying the many algorithms in current use. To this end, we have taken an important step toward unifying the great number of publications on the topic by demonstrating very general, deterministic conditions under which identification of parameters in this parsimonious model is not only possible but also robust to the inevitable uncertainty permeating measurement and model choice.

@@

An immediate application of our theoretical work is Prop.~\ref{neuroprop}, which certifies the validity of a theory of neurally-plausible bottleneck communication in the brain: that sparse coding in a compressed space of neural activity can recover sparse codes sent through a randomly-constructed (but unknown) noisy wiring bottleneck \cite{Isely10}.\footnote{We refer the reader to \cite{ganguli2012compressed} for more on interactions between dictionary learning and neuroscience.}

@@

In fact, uniqueness guarantees with minimal assumptions apply to all areas of data science and engineering that utilize learned sparse structure. For example, several groups have applied compressed sensing to signal processing tasks: MRI analysis \cite{lustig2008compressed}, image compression \cite{Duarte08}, and, more recently, the design of an ultrafast camera \cite{Gao14}. Given such effective uses of compressed sensing, it is only a matter of time before these systems incorporate dictionary learning to encode and decode signals (e.g., in a device that learns structure from motion \cite{kong2016prior}), just as scientists have used it to make sense of their data \cite{jung2001imaging, Agarwal14, lee2016sparse, wu2016stability}. Assurances such as those offered by our theorems certify that different devices (with different initialization, etc.) will learn equivalent representations given enough data from statistically identical systems.\footnote{To contrast with the current hot topic of ``Deep Learning'', there are few such uniqueness guarantees for these models of data; moreover, even small noise can dramatically alter their output \cite{goodfellow2014explaining}.} Indeed, it seems a main reason for the sustained interest in dictionary learning as an unsupervised method for data analysis is the assumed well-posedness of parameter identification in the sparse coding model, confirmation of which forms the core of our theoretical findings.

@@

%Since its inception some twenty years ago, sparse coding has become a standard tool in signal analysis, yielding myriad insights into the structure of natural signals across a variety of scientific domains. In this work,
